# Factors

Character vectors can be easily converted to factors via `as.factor()`:

```{r, size = "tiny", R.options = list(width = 100)}
degree = c("college", "PhD", "high school", "college", "PhD")
print(degree)
degree = as.factor(degree)
print(degree)
```

- Nominal and ordinal vectors should be converted to factors.

- A factor is basically an integer vector, where each number has a corresponding "label". 

<!-- - Often a vector is a character but only finitely many different values are  -->
<!-- allowed. -->
<!-- - Indexing factors is identical to vectors. -->

---

# Factors

The levels of a factor can be changed or reordered after it is created:
```{r, size = "tiny"}
print(degree)
levels(degree)[3] = "doctoral"
print(degree)

new.lvls = levels(degree)[c(2,1,3)]
print(new.lvls)
degree = factor(degree, levels = new.lvls)
print(degree)
```

<!-- ## Factors  -->

<!-- - In this case, an ordered factor makes sense (variables can be ordered by education level).  -->
<!-- This can be achieved by setting the argument `ordered = TRUE` -->

<!-- ```{r} -->
<!-- degree = factor(degree, ordered = TRUE) -->
<!-- degree[degree >= "college"] -->
<!-- ``` -->

---

# Factors

The factor levels can be extracted as a factor variable or as a character vector:

```{r}
unique(degree)
levels(degree)
```

- `unique()` keeps the factor structure.
- `levels()` returns the factor levels as a character vector.

---
